---
title: "stacked bar charts"
output: html_document
---

One of the first things that comes to mind when presenting metabarcoding data is the ever-popular stacked bar chart. Again, there are many packages that can be used to draw stacked bar charts (see for example [MicrobiotaProcess](https://github.com/YuLab-SMU/MicrobiotaProcess/)), but let's do it manually this time instead.

Let's load our latest Phyloseq object and librariries. 
```{r, message=FALSE,echo=TRUE}
library("phyloseq")
library("tidyverse")

WorkingPS<-readRDS("Processed_data/combinedPS_clean_pruned_filtered.rds");WorkingPS



# Because The Nanopore data set also contains fungi, lets keep only the Oomycota Phylum. 
WorkingPS<-subset_taxa(WorkingPS, Phylum=="Oomycota");WorkingPS

```

Here is a function borrowed from the MicrobiotaProcess package that will be helpful later when plotting our bar chart. 

```{r, message=FALSE,echo=TRUE}
get_cols <- function (n){
  col <- c("#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3",
           "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd",
           "#ccebc5", "#ffed6f")
  
  col2 <- c("#1f78e4", "#ffff33", "#c2a5cf", "#ff7f00", "#810f7c",
            "#a6cee3", "#006d2c", "#4d4d4d", "#8c510a", "#d73027",
            "#78c679", "#7f0000", "#41b6c4", "#e7298a", "#54278f")
  
  col3 <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
            "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a",
            "#ffff99", "#b15928")
  colorRampPalette(col2)(n)
}
```

The phyloseq psmelt function converts our phyloseq object into a data frame. 

```{r}
WorkingPS_melt <- psmelt(WorkingPS)

head(WorkingPS_melt)
```


```{r}
genus_abundance <- WorkingPS_melt %>%
  group_by(seq_tech, Week) %>%
  mutate(rel_abund = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  select(-Abundance) %>%
  pivot_longer(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
               names_to="level",
               values_to="taxon");head(genus_abundance)
```

```{r}
genus_abundance %>%
  filter(level=="Genus") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(rel_abund = sum(rel_abund), .groups="drop") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(mean_rel_abund=mean(rel_abund), .groups="drop") %>% 
  ggplot(aes(x=Week, y=mean_rel_abund, fill=taxon))+
  geom_col()+
  facet_wrap(vars(seq_tech), nrow = 3)+
  labs(x="Week",
       y="Relative abundance (%)")+
  theme(legend.text = element_text(face="italic"))+
  scale_fill_manual(values=get_cols(29))+ # This line calls the "get_cols">" function above
  guides(fill= guide_legend(keywidth = 0.6, 
                            keyheight = 0.7, 
                            ncol=1))+
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = 0.5, 
                                   hjust=0.4))
```





```{r}
predefined_species = c("Peronospora manshurica",
                     "Peronospora variabilis",
                      "Bremia elliptica",
                      "Hyaloperonospora nesliae",
                      "Peronospora destructor",
                      "Hyaloperonospora camelinae",
                      "Peronospora aparines",
                      "Peronospora erucastri",
                      "Albugo candida",
                      "Plasmopara viticola",
                      "Bremia lactucae",
                      "Hyaloperonospora brassicae",
                      "Pseudoperonospora cubensis"
);predefined_species
```


```{r}
WorkingPS_melt2 <-  WorkingPS_melt %>%
  mutate(Species = case_when(
    Species %in% 
      predefined_species ~ Species,  # Keep valid species as is
    TRUE ~ "Other"  # Replace other species with "other"
  ))

head(WorkingPS_melt2)
```




```{r}
genus_abundance2 <- WorkingPS_melt2 %>%
  group_by(seq_tech, Week) %>%
  mutate(rel_abund = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  select(-Abundance) %>%
  pivot_longer(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
               names_to="level",
               values_to="taxon")%>%
  mutate(taxon = factor(taxon,
                       levels=c("Peronospora manshurica",
                                "Peronospora variabilis",
                                "Peronospora destructor",                       
                                "Peronospora aparines",
                                "Peronospora erucastri",
                                "Bremia elliptica",
                                "Bremia lactucae",
                                "Albugo candida",
                                "Hyaloperonospora nesliae",
                                "Hyaloperonospora brassicae",
                                "Hyaloperonospora camelinae",
                                "Plasmopara viticola",
                                "Pseudoperonospora cubensis",
                                "Other")))
```



```{r}
genus_abundance2 %>%
  filter(level=="Species") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(rel_abund = sum(rel_abund), .groups="drop") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(mean_rel_abund=mean(rel_abund), .groups="drop") %>% 
  ggplot(aes(x=Week, y=mean_rel_abund, fill=taxon))+
  geom_col()+
  facet_wrap(vars(seq_tech), nrow = 3)+
  labs(x="Week",
       y="Relative abundance (%)")+
  theme(legend.text = element_text(face="italic"))+
  scale_fill_manual(values=get_cols(29))+ # This line calls the "get_cols">" function above
  guides(fill= guide_legend(keywidth = 0.6, 
                            keyheight = 0.7, 
                            ncol=1))+
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = 0.5, 
                                   hjust=0.4))
```
