---
title: "stacked bar charts"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

One of the first things that comes to mind when presenting metabarcoding data is the ever-popular stacked bar chart. Again, there are many packages that can be used to draw stacked bar charts (see for example [MicrobiotaProcess](https://github.com/YuLab-SMU/MicrobiotaProcess)), but let's do it manually this time instead.

Let's load our latest Phyloseq object and librariries. 
```{r, message=FALSE,echo=TRUE}
library("phyloseq")
library("tidyverse")

WorkingPS<-readRDS("Processed_data/combinedPS_clean_pruned_filtered.rds");WorkingPS



# Because The Nanopore data set also contains fungi, lets keep only the Oomycota Phylum. 
WorkingPS<-subset_taxa(WorkingPS, Phylum=="Oomycota");WorkingPS

saveRDS(WorkingPS, file= "Processed_data/WorkingPS.rds")
```

Here is a function borrowed from the MicrobiotaProcess package that will be helpful later when plotting our bar chart. 

```{r, message=FALSE,echo=TRUE}
get_cols <- function (n){
  col <- c("#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3",
           "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd",
           "#ccebc5", "#ffed6f")
  
  col2 <- c("#1f78e4", "#ffff33", "#c2a5cf", "#ff7f00", "#810f7c",
            "#a6cee3", "#006d2c", "#4d4d4d", "#8c510a", "#d73027",
            "#78c679", "#7f0000", "#41b6c4", "#e7298a", "#54278f")
  
  col3 <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
            "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a",
            "#ffff99", "#b15928")
  colorRampPalette(col2)(n)
}
```

Summarize the data 
----

The phyloseq psmelt function converts our phyloseq object into a data frame. 

```{r}
WorkingPS_melt <- psmelt(WorkingPS)

head(WorkingPS_melt)
```

Using the dplyr package (included in the tidyverse package) we create a tibble to summarize abundance and use the pivot_longer function to transpose the taxonomy from column to row.  


```{r}
genus_abundance <- WorkingPS_melt %>%
  group_by(seq_tech, Week) %>%
  mutate(rel_abund = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  select(-Abundance) %>%
  pivot_longer(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
               names_to="level",
               values_to="taxon")

str(genus_abundance)
```

Genus bar chart 
----

We then summarize the data, by sequencing technology and by week, at the genus level and input them into ggplot2 to obtain our bar chart with weeks on the x-axis and relative abundance on the y-axis, with facets for sequencing technologies. 

```{r}
genus_abundance %>%
  filter(level=="Genus") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(rel_abund = sum(rel_abund), .groups="drop") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(mean_rel_abund=mean(rel_abund), .groups="drop") %>% 
  ggplot(aes(x=Week, y=mean_rel_abund, fill=taxon))+
  geom_col()+
  facet_wrap(vars(seq_tech), nrow = 3)+
  labs(x="Week",
       y="Relative abundance (%)")+
  theme(legend.text = element_text(face="italic"))+
  scale_fill_manual(values=get_cols(29))+ # This line calls the "get_cols">" function above
  guides(fill= guide_legend(keywidth = 0.6, 
                            keyheight = 0.7, 
                            ncol=1))+
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = 0.5, 
                                   hjust=0.4))
```

Species bar chart
---- 

The first bar chart shows that the profiles are fairly similar between the three technologies, although some genera appear to be over- or under-represented with certain technologies. We would now like to examine the distribution of species as a function of time and sequencing technology. However, with over 80 species, it is difficult to get a clear representation of species distribution with a stacked bar chart, and we need to reduce the number of species to be plotted. 

To do that let's start by creating a list of predefined species that we want to be plotted in our bar chart: 


```{r}
predefined_species = c("Albugo candida",
                       "Bremia lactucae cluster",
                       "Hyaloperonospora nesliae",
                       "Hyaloperonospora brassicae",
                       "Hyaloperonospora camelinae",
                       "Peronospora manshurica",
                       "Peronospora variabilis",
                       "Peronospora destructor",                       
                       "Peronospora aparines",
                       "Peronospora erucastri",
                       "Plasmopara viticola",
                       "Pseudoperonospora cubensis",
)
```


And replace anything not in our list with "Other":

```{r}
WorkingPS_melt2 <-  WorkingPS_melt %>%
  mutate(Species = case_when(
    Species %in% 
      predefined_species ~ Species,  # Keep valid species as is
    TRUE ~ "Other"  # Replace other species with "other"
  ))

head(WorkingPS_melt2)
```

Generate a second tibble with using this predefined list of species: 


```{r}
genus_abundance2 <- WorkingPS_melt2 %>%
  group_by(seq_tech, Week) %>%
  mutate(rel_abund = Abundance / sum(Abundance)) %>%
  ungroup() %>%
  select(-Abundance) %>%
  pivot_longer(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
               names_to="level",
               values_to="taxon")%>%
  mutate(taxon = factor(taxon,
                       levels=c("Albugo candida",
                       "Bremia lactucae cluster",
                       "Hyaloperonospora nesliae",
                       "Hyaloperonospora brassicae",
                       "Hyaloperonospora camelinae",
                       "Peronospora manshurica",
                       "Peronospora variabilis",
                       "Peronospora destructor",                       
                       "Peronospora aparines",
                       "Peronospora erucastri",
                       "Plasmopara viticola",
                       "Pseudoperonospora cubensis",
                       "Other")))
```

summarize the data at the species level and pipe that into ggplot2 to get our bar chart at the species level: 

```{r}
genus_abundance2 %>%
  filter(level=="Species") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(rel_abund = sum(rel_abund), .groups="drop") %>% 
  group_by(seq_tech, Week, taxon) %>% 
  summarize(mean_rel_abund=mean(rel_abund), .groups="drop") %>% 
  ggplot(aes(x=Week, y=mean_rel_abund, fill=taxon))+
  geom_col()+
  facet_wrap(vars(seq_tech), nrow = 3)+
  labs(x="Week",
       y="Relative abundance (%)")+
  theme(legend.text = element_text(face="italic"))+
  scale_fill_manual(values=get_cols(14))+ # This line calls the "get_cols">" function above
  guides(fill= guide_legend(keywidth = 0.6, 
                            keyheight = 0.7, 
                            ncol=1))+
  theme(axis.text.x = element_text(angle = 60, 
                                   vjust = 0.5, 
                                   hjust=0.4))
```

Again we can see some similitude and differences between sequencing technologies, now lets see how different they are. 
