---
title: "Temporal diversity"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

To begin exploring the distribution of species across the data set, the use of heat maps provides an at-a-glance overview of distribution at different levels.

Several packages can be used to create heat maps, including Phyloseq, Microbiome, MicrobiotaProcess, microeco and others. Here we'll use Ampvis2 to generate our graphs. 

```{r, message=FALSE,echo=TRUE}
library("ampvis2")
library("ggpubr")

WorkingPS_glom<-readRDS("Processed_data/WorkingPS_glom.rds")
```

Heat map by week and technologies  
---- 

First we need to convert our phyloseq object into an Ampvis2 object, then generate the heat map with the weeks (times) in x and the sequencing technology as facets. 

```{r, fig.align="center", echo = TRUE, fig.width = 14, fig.height= 6}
# Convert the PS object to ampvis 
WorkingPS_amp <- amp_load(WorkingPS_glom)

# Plot the heat map
amp_heatmap(
  WorkingPS_amp,
  group_by = "Week2",
  facet_by = "seq_tech",
  normalise = TRUE,
  tax_aggregate = "Species",
  tax_add = NULL,
  tax_show = 15,
  showRemainingTaxa = FALSE,
  tax_class = NULL,
  tax_empty = "remove",
  plot_values = TRUE,
  plot_values_size = 3,
  plot_legendbreaks = NULL,
  plot_colorscale = "log10",
  plot_na = FALSE,
  measure = mean,
  min_abundance = 10,
  max_abundance = NULL,
  round = 1)+
  theme(axis.text.x = element_text(angle = 90, 
                                   size=8, 
                                   vjust = 0),
        axis.text.y = element_text(size=12, 
                                   face="italic"),
        legend.position="bottom")
```

Heat map by week and site   
----

Let's start again by putting sites in the place of sequencing technologies. 

```{r, fig.align="center", echo = TRUE, fig.width = 16, fig.height= 8}
# Plot the heat map
amp_heatmap(
  WorkingPS_amp,
  group_by = "Week2",
  facet_by = "SiteID",
  normalise = TRUE,
  tax_aggregate = "Species",
  tax_add = NULL,
  tax_show = 15,
  showRemainingTaxa = FALSE,
  tax_class = NULL,
  tax_empty = "remove",
  plot_values = TRUE,
  plot_values_size = 3,
  plot_legendbreaks = NULL,
  plot_colorscale = "log10",
  plot_na = FALSE,
  measure = mean,
  min_abundance = 10,
  max_abundance = NULL,
  round = 1)+
  theme(axis.text.x = element_text(angle = 90, 
                                   size=8, 
                                   vjust = 0),
        axis.text.y = element_text(size=12, 
                                   face="italic"),
        legend.position="bottom")
```

Heat map for _Peronospora_ 
----

That's a lot of information, so let's try to extract genera of interest to reduce the number of taxa to display.

```{r, fig.align="center", echo = TRUE, fig.width = 14, fig.height= 6}
PeronosporaPS<-subset_taxa(WorkingPS_glom, Genus=="Peronospora");PeronosporaPS


Peronospora_amp <- amp_load(PeronosporaPS)

amp_heatmap(
  Peronospora_amp,
  group_by = "Week2",
  facet_by = "SiteID",
  normalise = TRUE,
  tax_aggregate = "Species",
  tax_add = NULL,
  tax_show = 12,
  showRemainingTaxa = FALSE,
  tax_class = NULL,
  tax_empty = "remove",
  plot_values = TRUE,
  plot_values_size = 2,
  plot_legendbreaks = NULL,
  plot_colorscale = "log10",
  plot_na = FALSE,
  measure = mean,
  min_abundance = 10,
  max_abundance = NULL,
  round = 1)+
  theme(axis.text.x = element_text(angle = 90, 
                                   size=8, 
                                   vjust = 0),
        axis.text.y = element_text(size=12, 
                                   face="italic"),
        legend.position="bottom")
```
